\documentclass[12pt,preprint]{aastex}
%\documentclass[iop]{emulateapj}
%\documentclass{article}
%\usepackage{draftwatermark}
\usepackage{paralist}
\shorttitle{Magnetized massive disks}
\shortauthors{M-K. Lin}

\usepackage{amsmath}
%\usepackage{url}
\usepackage{bm}
%\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
\newcommand{\p}{\partial}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\sbar}{\bar{\sigma}} 
\newcommand{\imgi}{\mathrm{i}} 
\newcommand{\ii}{\mathrm{i}}
\newcommand{\C}{\mathcal{C}^\lambda}
\newcommand{\wop}[1]{\mathcal{V}^{(#1)}}
\newcommand{\avg}[1]{\langle{#1}\rangle}
\newcommand{\zhat}{\hat{z}}
\newcommand{\rhat}{\hat{r}}
\newcommand{\he}{\mathcal{H}}
\newcommand{\amp}{\mathcal{A}}
\newcommand{\dd}{\delta}
\newcommand{\real}{\operatorname{Re}}
\newcommand{\imag}{\operatorname{Im}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\zmax}{Z_s}
\newcommand{\wbar}{\tilde{W}}
\newcommand{\qbar}{\tilde{Q}}
\newcommand{\dbigH}{\frac{H^\prime}{H}}
\newcommand{\Rp}{\frac{\rho_0^\prime}{\rho_0}}
\newcommand{\Fp}{\frac{g^\prime}{g}}
\newcommand{\Rpp}{\frac{\rho_0^{\prime\prime}}{\rho_0}}
\newcommand{\Fpp}{\frac{g^{\prime\prime}}{g}}
\newcommand{\ils}{L_s^{-1}}
\newcommand{\ilp}{L_p^{-1}}
\newcommand{\ihp}{H_p^{-1}}
\newcommand{\ihs}{H_s^{-1}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\csmid}{c_{s0}} 
\newcommand{\hiso}{h_\mathrm{iso}}
\newcommand{\Hiso}{H_\mathrm{iso}}
\newcommand{\zeus}{\texttt{ZEUS-MP }}
\newcommand{\rsph}{r_\mathrm{sph}}
\newcommand{\dvx}{\dd v_x}
\newcommand{\dvy}{\dd v_y}
\newcommand{\dvz}{\dd v_z}
\newcommand{\dbx}{\dd B_x}
\newcommand{\dby}{\dd B_y}
\newcommand{\dbz}{\dd B_z}
\newcommand{\w}{ \widetilde{W}}
%\newcommand{\sbar}{\bar{\sigma}}
\newcommand{\dphi}{\dd \Phi}
\defcitealias{lin12}{L12} 

\begin{document}

\title{Marginal local stability analysis of self-gravitating Keplerian
  disks with polytropic index $n=1$ and a vertical magnetic field}   


\author{Min-Kai Lin}
\affil{Canadian Institute for Theoretical Astrophysics,
  60 St. George Street, Toronto, ON, M5S 3H8, Canada}
\email{mklin924@cita.utoronto.ca}

\section{Problem setup}
Consider axisymmetric perturbations in the shearing box
framework. Equation of state is $P\propto \rho^2$. Then the 
square of the sound-speed $c_s^2\equiv dP/d\rho \propto \rho$. For a
Keplerian disk, the equilibrium density field is
\begin{align}
\hat{\rho}(z)\equiv \frac{\rho}{\rho_0} = (1+Q)\cos{az} - Q, 
\end{align}
where $\rho_0\equiv\rho(0)$, $Q\equiv \Omega^2 / 4\pi G\rho_0$ and
$aH\equiv \arccos{Q/(1+Q)}$ so that $\rho(H) = 0$, i.e. $H$ is the
disk thickness. For the magnetic field, assume ideal 
MHD. The initial field is vertical.      

\subsection{Normalization}
Lengths are normalized by the disk thickness $H$. Velocities
normalized by the midplane sound-speed $c_{s0}$. Magnetic field
normalized by initial field $B_z$. Frequencies normalized by the
rotation frequency $\Omega$. Potential and perturbed enthalpy normalized by
$c_{s0}^2$. Density normalized by $\rho_0$. Note that for a polytropic
disk with $n=1$, $c_s^2=\rho$ in normalized 
units.  


\section{Governing equations}
The non-dimensional linearized equations are listed here. The
continuity equation is 
\begin{align}
  \frac{\imgi\sigma\rho}{c_s^2}W + \imgi  f k_x \rho \dvx +
  f\left(\rho\dvz\right)^\prime = 0, \label{cont}   
\end{align}
where $W = c_s^2\delta\rho/\rho$ is the enthalpy perturbation, and 
$f= c_{s0}/H\Omega$. The eigenfrequency is $\sigma$ such that
perturbations have time dependence $\exp{\imgi\sigma t}$. 


The equations of motion are
\begin{align}
  \imgi\sigma \dvx - 2 \dvy  &= - \imgi
  f k_x\w + fv_A^2 \left(\dbx^\prime - \imgi 
  k_x\dbz\right),\\
  \imgi\sigma\dvy + 
  \frac{\kappa^2}{2}\dvx &= fv_A^2\dby^\prime\label{lin_vy},\\ 
  \imgi\sigma\dvz & = -f \w^\prime\label{lin_vz}, 
\end{align}
where $\w \equiv W + \delta\Phi$, and the square of the Alfven speed
$v_A^2= (\beta\rho)^{-1}$ in normalized units. 

The components of the induction equation are 
\begin{align}
&  \imgi\sigma\dbx = f \dvx^\prime\label{induct_x}\\
&\imgi \sigma \dby = f\dvy^\prime - S\dbx, \label{induct_y}\\
& \imgi\sigma \dbz = -\imgi  f k_x \dvx, \label{induct_vert}
%& \imgi\bar{\sigma}\dbz = -\imgi  k_x \dvx  -\imgi k_x\eta \dbx^{\prime}, \label{induct_vert}
\end{align}    
and the divergence-free condition is $\imgi k_x\dbx +
\dbz^\prime=0$. 

The linearized Poisson equation is 
\begin{align}
  \dphi^{\prime\prime} - k_x^2\dphi =
  \frac{\Omega^2\rho}{c_s^2Qf^2}W.  \label{lin_poisson} 
\end{align}

Using the induction equation to eliminate the magnetic field from the
horizontal equations of motion gives
\begin{align}
f^2v_A^2\left(\dvx^{\prime\prime} - k_x^2\dvx\right) + \sigma^2\dvx +
2\imgi\sigma\dvy + \sigma f k_x \w &= 0,\label{vx_new}\\
f^2v_A^2\left(\sigma\dvy^{\prime\prime} + \imgi S
\dvx^{\prime\prime}\right) + \sigma^3\dvy - \frac{\imgi\sigma^2\kappa^2}{2}\dvx&=0.\label{vy_new}
\end{align}


\subsection{Boundary conditions}
Even symmetry at the mid-plane,
\begin{align}
  \dvx^\prime(0) = \dvy^\prime(0)=W^\prime(0) = \dphi^\prime(0) = 0.
\end{align}
Assume a vertical field at the disk surface, $\delta B_x(1) = \delta
B_y(1) = 0$, or 
\begin{align}
  \dvx^\prime(1) = \dvy^\prime(1) = 0. 
\end{align}
Impose no vertical velocity at the disk surface, $\dvz(1) = 0$, or 
\begin{align}
  W^\prime(1) + \dphi^\prime(1) = 0. 
\end{align}
The boundary condition for the potential is
\begin{align}
  \dphi^\prime (1) + k_x \dphi(1) = 0.  \label{pot_cond}
\end{align}

\section{Marginal stability} 
Consider the problem in the limit $\sigma \to 0$. Assume perturbations
have power series solutions in $\sigma$, i.e. 
\begin{align}
  \dvx &= \sum_{n=0}^\infty u_n(z)\sigma^n,\\
  \dvy &= \sum_{n=0}^\infty v_n(z)\sigma^n,\\
  W    &= \sum_{n=0}^\infty w_n(z)\sigma^n,\\
  \dphi&= \sum_{n=0}^\infty \phi_n(z)\sigma^n, 
\end{align}
and $\w = \sum_{n=0}^\infty \tilde{w}_n \sigma^n$ with $\tilde{w}_n =
w_n + \phi_n$. 

\subsection{The Poisson equation} 
Assuming a similar expansion as above for the vertical velocity, at
$O(\sigma^0)$ the vertical equation of motion (Eq. \ref{lin_vz}) reads
\begin{align}
  \tilde{w}^\prime_0 = 0. 
\end{align}
This implies
\begin{align}
  w_0(z) + \phi_0(z) \equiv \tilde{w} _0 = \mathrm{constant}. 
\end{align}
The Poisson equation reads, after setting $\dphi = \w - W$ and
$c_s^2=\rho$, 
\begin{align}
  \left(\tilde{w}_n-w_n\right)^{\prime\prime} - k_x^2\left(\tilde{w}_n - w_n\right) =
  \frac{1}{f^2Q} w_n. 
\end{align}
For $n=0$ this is
\begin{align}
  w_0^{\prime\prime} + \underbrace{\left(\frac{1}{f^2Q} -
    k_x^2\right)}_{\tilde{k}_x^2}w_0 = -k_x^2 \tilde{w}_0, 
\end{align}
and the RHS is a constant. The boundary conditions are 
\begin{align}
  &w_0^\prime(0) = 0,\\
  &w_0^\prime(1) + k_x w_0(1) = k_x \tilde{w}_0, \label{poisson_zero}
\end{align} 
where the condition at the disk surface is implied by the potential
condition (Eq. \ref{pot_cond}). The solution to Eq. \ref{poisson_zero}
is then
\begin{align}
  w_0(z) = \frac{\tilde{w}_0\left(1 +
    k_x^2/\tilde{k}_x^2\right)}{\cos{\tilde{k}_x} -
    \left(\tilde{k}_x/k_x\right)\sin{\tilde{k}_x}}\cos{\tilde{k}_xz} -
  \frac{k_x^2}{\tilde{k}_x^2}\tilde{w}_0. \label{w0_sol}  
\end{align}

\subsection{The continuity equation} 
Integrating the continuity equation gives
\begin{align}
  \sigma \int_0^1 W dz + f k_x \int_0^1\rho \dvx = 0,  \label{int_cont0}
\end{align}
where $c_s^2 = \rho$ has been used, as well as the boundary condition
$\dvz(0)=\dvz(1)=0$. At $O(\sigma^1)$ this equation reads
\begin{align}
  \int_0^1w_0 dz + f k_x \int_0^1\rho u_1 dz = 0.
\end{align}
Inserting the above solution for $w_0$ (Eq. \ref{w0_sol}), 
\begin{align}
   \frac{\tilde{w}_0\left(1 +
    k_x^2/\tilde{k}_x^2\right)}{\left[\cos{\tilde{k}_x} -
    \left(\tilde{k}_x/k_x\right)\sin{\tilde{k}_x}\right]}\frac{\sin{\tilde{k}_x}}{\tilde{k}_x} -
  \frac{k_x^2}{\tilde{k}_x^2}\tilde{w}_0 + fk_x \int_0^1\rho u_1 dz.
  =0 \label{int_cont} 
\end{align}
Note that at $O(\sigma^0)$, Eq. \ref{int_cont0} is satisfied by
$u_0=0$, which is consistent with the horizontal equations of motion
discussed next. 


\subsection{Horizontal equations of motion}
\noindent Eq. \ref{vx_new} at $O(\sigma^0)$,
\begin{align}
  u_0^{\prime\prime} - k_x^2u_0 = 0.
\end{align}
Eq. \ref{vx_new} at $O(\sigma^1)$,
\begin{align}
  f^2v_A^2\left(u_1^{\prime\prime} - k_x^2u_1\right)  + 2 \imgi v_0 +
  f k_x \tilde{w}_0= 0. \label{vx_o1} 
\end{align}
Eq. \ref{vx_new} at $O(\sigma^2)$,
\begin{align}
  f^2v_A^2\left(u_2^{\prime\prime} - k_x^2u_2\right)  + u_0 + 2 \imgi v_1 +
  f k_x \tilde{w}_1= 0.   
\end{align}
Eq. \ref{vy_new} at $O(\sigma^0)$,
\begin{align}
  u_0^{\prime\prime} = 0.
\end{align}
Eq. \ref{vy_new} at $O(\sigma^1)$,
\begin{align}
  v_0^{\prime\prime} + \imgi S u_1^{\prime\prime} = 0.\label{v0pp}
\end{align}
Eq. \ref{vy_new} at $O(\sigma^2)$,
\begin{align}
  f^2v_A^2\left(v_1^{\prime\prime} + \imgi S u_2^{\prime\prime}\right)
  - \frac{\imgi\kappa^2}{2}u_0= 0.  
\end{align}
Eq. \ref{vy_new} at $O(\sigma^3)$,
\begin{align}
  f^2v_A^2\left(v_2^{\prime\prime} + \imgi S u_3^{\prime\prime}\right)
  + v_0 - \frac{\imgi\kappa^2}{2}u_1= 0.  \label{vy_o3} 
\end{align}

\subsubsection{Relation between $v_0$ and $u_1$}
Integrating Eq. \ref{v0pp} and imposing the midplane symmetry
condition gives
\begin{align}
  v_0(z) = K -\imgi S u_1(z),  
\end{align}
where $K$ is a constant. Multiplying this equation by $\rho$ and
integrating, 
\begin{align}
  \int_0^1\rho v_0 dz = \frac{K\Sigma}{2} - \imgi S \int_0^1\rho u_1 dz, 
\end{align}
where $\Sigma = \int_{-1}^1\rho dz$ is the dimensionless column
density. 

Recalling that $v_A^2\propto 1/\rho$, multiply Eq. \ref{vy_o3} by
$\rho$ and integrating,
\begin{align}
  \int_0^1\rho v_0 dz = \frac{\imgi\kappa^2}{2} \int_0^1\rho u_1 dz, 
\end{align} 
where the midplane symmetry condition was used. Hence
\begin{align}
  K = \frac{\imgi}{\Sigma}\left(2S + \kappa^2\right)\int_0^1\rho u_1
  dz. \label{K_def}
\end{align}

\subsubsection{Solution for $u_1$} 
Inserting $v_0 = K - \imgi S u_1$ into Eq. \ref{vx_o1} and
rearranging, 
\begin{align}
  u_1^{\prime\prime}(z) - k_x^2u_1(z) +
  \frac{2S\beta}{f^2}\rho(z)u_1(z) =
  -\frac{\beta}{f^2}\left(fk_x\tilde{w}_0 + 2\imgi K\right)\rho(z)\equiv
  g(z).  
\end{align}
Inserting the expression for the density field,
\begin{align}
  u_1^{\prime\prime}(z) - \left( k_x^2 + \frac{2\beta S
    Q}{f^2}\right)u_1(z) + 2\frac{\beta S(1+Q)}{f^2}\cos{(az)}
  u_1(z) = g(z). 
\end{align}
Let
\begin{align}
  az &= 2X,\\
  k_1 & =  -\frac{4}{a^2}\left( k_x^2 + \frac{2\beta S
    Q}{f^2}\right),\\
  k_2 & = -\frac{4}{a^2}\frac{\beta S(1+Q)}{f^2},\\
  \bar{g} &= \frac{4}{a^2} g.
\end{align}
Then
\begin{align}
  u_1^{\prime\prime}(X) + \left[k_1 - 2k_2\cos{\left(2X\right)} \right]u_1(X) =
  \bar{g}(X). \label{u1_eq} 
\end{align}
The homogeneous part of Eq. \ref{u1_eq} is Mathieu's differential
equation and is solved by Mathieu functions. We define the
complementary functions
\begin{align}
  y_a(X) &= M_c(k_1, k_2, X),\\
  y_b(X) &= M_c(k_1, k_2, X)\frac{M_{s1}^\prime}{M_{c1}^\prime} -
  M_s(k_1, k_2, X),  
\end{align}
where $M_c$ and $M_s$ are even and odd Mathieu functions,
respectively, and subscript $1$ denotes evaluation at $X=a/2$
($z=1$). Note that we have chosen $y_a$ to satisfy the symmetry
condition at $z=0$ and $y_b$ to satisfy the upper disk boundary
condition. We will omit writing the dependence on $k_1$ and $k_2$ from
now on. 

The inhomogeneous problem, Eq. \ref{u1_eq}, is solved with the
variation of parameters,
\begin{align}
  u_1(X) =
  y_b(X)\int_0^X\frac{y_a(\xi)\bar{g}(\xi)}{\mathcal{W}}d\xi +
  y_a(X) \int_X^{a/2}\frac{y_b(\xi)\bar{g}(\xi)}{\mathcal{W}}d\xi, 
\end{align} 
where the Wronskian is defined as 
\begin{align}
  \mathcal{W} \equiv y_ay_b^\prime - y_by_a^\prime,
\end{align}
and is a constant for the choice of $y_{a,b}$ above. Inserting the
definition for $\bar{g}$ gives
\begin{align}
  u_1(X) = -\frac{4}{a^2}\frac{\beta}{f^2\mathcal{W}}\left(fk_x\tilde{w}_0 +
  2\imgi K \right)\left[
    y_b(X)\int_0^X y_a(\xi)\rho(\xi) d\xi +
    y_a(X) \int_X^{a/2}y_b(\xi)\rho(\xi)d\xi
    \right],\label{u1_explicit}
\end{align} 
and note that $\rho(\xi) = (1+Q)\cos{2\xi} - Q$. 

Next, define
\begin{align}
\Theta \equiv \frac{1}{\mathcal{W}}\int_0^{a/2}\rho(X)y_b(X)dX\left[\int_0^X y_a(\xi)\rho(\xi) d\xi\right] +
    \int_0^{a/2}\rho(X)y_a(X)dX
    \left[\int_X^{a/2}y_b(\xi)\rho(\xi)d\xi\right], 
\end{align}
so that multiplying Eq. \ref{u1_explicit} by $\rho$ and integrating it
gives 
\begin{align}
\int_0^{a/2}\rho(X)u_1(X)dX &= -\frac{4\beta\Theta}{a^2 f^2
  }\left(fk_x\tilde{w}_0 + 2\imgi K\right),\notag\\
&= -\frac{4\beta\Theta}{a^2 f^2
}\left[fk_x\tilde{w}_0 - \frac{2(2S +
    \kappa^2)}{\Sigma}\int_0^{1}\rho(z)u_1(z)dz\right], 
\end{align}
where the expression for $K$ is used (Eq. \ref{K_def}). Inserting $X = az/2$ on the LHS gives
\begin{align}
  \frac{a^2f^2}{4\beta\Theta}\frac{a}{2}\int_0^1\rho(z)u_1(z)dz = -
  fk_x\tilde{w}_0 +
  \frac{2(2S+\kappa^2)}{\Sigma}\int_0^1\rho(z)u_1(z)dz, 
\end{align}
or
\begin{align}
  fk_x\tilde{w}_0 = \left[\frac{2(2S + \kappa^2)}{\Sigma} -
    \frac{a^3f^2}{8\beta\Theta}\right]\int_0^1\rho(z)u_1(z)dz. \label{int_horizontal}
\end{align}

\subsection{Condition for marginal stability}
Eliminating the integrals between Eq. \ref{int_cont} and
Eq. \ref{int_horizontal} gives
\begin{align}
   \frac{\left(1 +
    k_x^2/\tilde{k}_x^2\right)}{\left[\cos{\tilde{k}_x} -
    \left(\tilde{k}_x/k_x\right)\sin{\tilde{k}_x}\right]}\frac{\sin{\tilde{k}_x}}{\tilde{k}_x} -
  \frac{k_x^2}{\tilde{k}_x^2} + \frac{f^2k_x^2
    }{\left[\frac{2(2S + \kappa^2)}{\Sigma} -
    \frac{a^3f^2}{8\beta\Theta}\right]} = 0.
\end{align}
Given $Q$ and $\beta$, this equation should be solved for $k_x$. 

%\begin{abstract}
%\end{abstract}
%\section{Introduction}
%\input{setup}
%\input{linear}
%\input{results}
%\input{results_gi}
%\appendix
%\input{appendix}
%\bibliographystyle{apj}
%\bibliography{ref}


\end{document}

